--------------------------------------------------------------------------------
title: "p8105_hw1_hmb2161"
author: Hannah Bowlin
output: github_document
--------------------------------------------------------------------------------

# Problem 0.1:

-Set up a public GitHub repo and local R project: p8105_hw1_hmb2161 (github link: https://github.com/hmbowlin/p8105_hw1_hmb2161) 
-Create a single .rmd file named p8105_hw1_hmb2161.rmd (in use) that goes to github document (linked and the output github has been changed from html above)
-Knit document to ensure work is reproducible:

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Problem 0.2:

Ensure solutions are correctly edited and styled based on the guidelines in Building Blocks.

# Problem 1:

I created a code chunk to answer Problem 1 of the HW1. This problem intends to look at variable types (numeric, character, and factor) and basic coercion.

**Part I: Creating dataframe for different vectors** 
```{r}
##I set the working library at tidyverse.
library(tidyverse)

## I set the seed to ensure I get the same numbers each time I run this code.
set.seed(100)

## I created a random sample of size 8 from the standard normal distribution, a logical vector showing if the elements in the sample are greater than zero (it will return "false" if it doesn't satisfy this condition), a character vector that is 8 elements long with three different factors apple, pear, and banana.
sample_number = rnorm(n = 8)
vec_logical = sample_number > 0
vec_character = character(length = 8)
vec_factor= factor(c("apple", "banana", "pear", "banana", "pear", "apple", "apple", "pear"))

## I will now take the mean of each variable in my data frame. 
mean(sample_number)
mean(vec_logical)
mean(vec_character)
mean(vec_factor)
```
-The mean for (sample_number) is 0.1256937. As sample_number is a random sample of numbers and I am able to take the mean of these numbers
-The mean for (vec_logical) is 0.625. As vec_logical is based on sample_number, it will run when its condition is satisfied. vec_logical represents the mean of sample_number when sample_number's elements are greater than 0.
-The mean for vec_character is not possible as R views it as non-numeric text.You could store numbers in a character vector, but R would not view them as numbers, but as numbers stored as text.
-The mean for vec_factor is not possible as R views it as non-numeric characters that are grouped into factor levels.

Part II: Converting variables between character and numeric format
```{r, results = hide}
as.numeric(vec_logical)
as.numeric(vec_character)
as.numeric(vec_factor)
```
I created a new code chunk to work on the command as.numeric. I changed the logical, character, and factor vectors to numeric. The sample_number vector was already numeric and doesn't need to be changed. 

The command made these character and factor vectors numeric so the logical vector now stores true and false as 0 and 1, the character vector stores the values as 8 numbers, and when I entered in factor levels it saves as 1, 2, 3, not apple, banana, pear.

This means that you can change character vectors into numeric form with the command as.numeric. I now know that I couldn't take the mean of these variables because they weren't stored as numbers, they were stored as text.

Part III: Manipulation of vectors and formats
```{r, results = hide}
#Changed vec_logical to numeric and multiplied it by random sample
as.numeric(vec_logical) * sample_number 

#Changed vec_logical to factor and multiplied it by the random sample
as.numeric(vec_factor) * sample_number

#Changed vec_logical to factor and converted result to numeric and then multipled it by the random sample
as.numeric(as.factor(vec_logical)) * sample_number
```

