##################################################################
# Sept 20, 2019
# 
# Hannah Bowlin
# 
# Script for P8105 HW 1
#################################################################

---
title: "p8105_hw1_hmb2161"
output:
  github_document: default
  pdf_document: default
---

Problem 0.1:

-Set up a public GitHub repo and local R project: p8105_hw1_hmb2161 (github link: https://github.com/hmbowlin/p8105_hw1_hmb2161) 
-Create a single .rmd file named p8105_hw1_hmb2161.rmd (in use) that goes to github document (linked and the output github has been changed from html above)
-Knit document to ensure work is reproducible:

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Problem 1:

I created a code chunk to answer Problem 1 of the HW1. This problem intends to look at how well I know variable types (numeric, character, and factor) and some coercion.


```{r}
##I set the working library at tidyverse.
library(tidyverse)

## I set the seed to ensure I get the same numbers each time I run this code.
set.seed(100)

## I created a random sample of size 8 from the standard normal distribution, a logical vector showing if the elements in the sample are greater than zero (it will return "false" if it doesn't satisfy this condition), a character vector that is 8 elements long with three different factors apple, pear, and banana.
sample_number = rnorm(n = 8)
vec_logical = sample_number > 0
vec_character = character(length = 8)
vec_factor= factor(c("apple", "banana", "pear", "banana", "pear", "apple", "apple", "pear"))

## I will now take the mean of each variable in my data frame. 
mean(sample_number)
mean(vec_logical)
mean(vec_character)
mean(vec_factor)

## The mean for (sample_number) is 0.1256937. As sample_number is a random sample of numbers and I am able to take the mean of these numbers
## The mean for (vec_logical) is 0.625. As vec_logical is based on sample_number, it will run when its condition is satisfied. vec_logical represents the mean of sample_number when sample_number's elements are greater than 0.
## The mean for vec_character is not possible as R views it as non-numeric text.You could store numbers in a character vector, but R would not view them as numbers, but as numbers stored as text.
## The mean for vec_factor is not possible as R views it as non-numeric characters that are grouped into factor levels.

## To work on the command as.numeric, I changed the logical, character, and factor vectors to numeric. The sample_number vector was already numeric and doesn't need to be changed. 
as.numeric(vec_logical)
as.numeric(vec_character)
as.numeric(vec_factor)

## The command made these character and factor vectors numeric so the logical vector now stores true and false as 0 and 1, the character vector stores the values as 8 numbers, and when I entered in factor levels it saves as 1, 2, 3, not apple, banana, pear.

## This means that you can change character vectors into numeric form with the command as.numeric. I now know that I couldn't take the mean of these variables because they weren't stored as numbers, they were stored as text.


```

